---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r loaddata}
# Load data into activity using read.csv
activity <- read.csv("activity.csv")

# Quick overview of activity
summary(activity)
```


## What is mean total number of steps taken per day?
``` {r}
# Calculate number of steps per day using tapply and removing NA's
TotalStepsPerDay <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)

# Plot resulting histogram
hist(TotalStepsPerDay, col = "red", breaks = 25,
     xlab = "Total steps per day",
     xlim = c(0, 25000))
```


We compute the mean and median values
``` {r}
mean(TotalStepsPerDay)
median(TotalStepsPerDay)
```
  
## What is the average daily activity pattern?
  
``` {r}
# Calculate average number of steps per interval using tapply and removing NA's
AverageStepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)

# Plot resulting Time Series plot
plot(names(AverageStepsPerInterval), AverageStepsPerInterval, 
     type = "l", 
     col = "blue", 
     main = "Average daily pattern", 
     xlab = "5-minute interval (time)", 
     ylab = "Average number of steps accross all days")
```

``` {r}
# Find max of AverageStepsPerInterval
ind_max <- AverageStepsPerInterval == max(AverageStepsPerInterval)
interval_max <- AverageStepsPerInterval[ind_max]
interval_max
```

The 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps is: `r names(interval_max)`
  
  
## Imputing missing values
First let's find the total number of missing values in the dataset:
``` {r} 
# Using summary
summary(activity$steps)

# Create logical vector of missing values
missing_values <- is.na(activity$steps)

# Compute length
nb_na <- length(activity$steps[missing_values])
nb_na
```
The number of missing values is: `r nb_na`  

We create a new dataset which will contain the same data than activity but with the missing data filled in
   
   

## Are there differences in activity patterns between weekdays and weekends?

``` {r} 

```

